---
title: "Data visualization"
subtitle: "POL051: The Scientific Study of Politics"  
author: 
  - "Prof. Juan Tellez"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    seal: true
    css: [xaringan-themer.css, "ath-slides.css"]
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
      navigation:
        scroll: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
library(flipbookr)
library(tidyverse)
style_mono_accent(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


```{r xaringanExtra-share-again, echo=FALSE}
xaringanExtra::use_share_again()
```

class: center middle main-title section-title-0

# Data Visualization

.class-info[

.medium[**Prof. Juan Tellez**]


.light.small[Andrew Young School of Policy Studies<br>
Georgia State University<br>
July 7, 2020]

]

---
# Hello

This is just a slide.
- whatever
* whatever

.box-1[Don't tell me what to do]

???

presenter notes.

---

# Code on one side and text on the other

.left-code[
```{r plot-label, eval=FALSE}
# code chunk here
ggplot(iris) + 
  aes(Sepal.Length, 
      Sepal.Width, 
      color = Species) + 
  geom_point()
```
]

.right-plot[
```{r plot-label-out, ref.label="plot-label", echo=FALSE}
```
]
---

# And here's some math

$$Y_{it} = \alpha + \beta X + \epsilon_{it}$$


---
class: inverse, middle, center

# First up: classic flipbook, using defaults to walk through code pipeline





<!-- get a flipbook version of the my_cars code chunk - pauses are where parentheses are balanced -->
---

`r chunk_reveal("my_cars")`

```{r my_cars, include = FALSE}
cars %>%
  filter(speed > 4) %>%
  ggplot() +
  aes(x = speed) + #BREAK
  aes(y = dist) + #BREAK
  geom_point(
    alpha = .8, 
    color = "blue" 
    ) + 
  aes(size = speed) #BREAK
```

