---
title: "DAGs"
citeproc: true
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-fullnote-bibliography-no-bib.csl
menu: 
  readings:
    parent: Readings
    weight: 8
type: docs
output:
  blogdown::html_page:
    toc: true
editor_options: 
  chunk_output_type: console
---


## Recommended

- work through the examples in the `ggdag` package vignette [ here](https://cran.r-project.org/web/packages/ggdag/vignettes/intro-to-dags.html).^[Don't just copy/paste the code; write it out yourself! Change the variables in the smoking/cancer example to something of interest to you.]

- <i class="fas fa-book"></i> [Directed Acyclical Graphs in Causal Inference: The Mixtape](https://mixtape.scunning.com)^[This reading is mathy-er than what we've covered so far. It also features code from STATA, which we don't use. Don't worry about that so much; we will focus on broad concepts/ideas.]


## Slides

### Tuesday

[Class slides here](/slides/09-dags.html)


```{r,echo=FALSE}
xaringanExtra::embed_xaringan(
   url = "/slides/09-dags.html",
   ratio = "16:9"
)
```


### Thursday


[Class slides here](/slides/10-controls-I.html)


```{r,echo=FALSE}
xaringanExtra::embed_xaringan(
   url = "/slides/10-controls-I.html",
   ratio = "16:9"
)
```