---
title: "Hell is other people (on planes)"
author: "Juan Tellez"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---


Let's look at some cringe-inducing survey data from [Five Thirty Eight](https://fivethirtyeight.com/features/airplane-etiquette-recline-seat/) on all the awkward things that people have to think about while flying. The survey captures 1040 responses and was administered on SurveyMonkey. 




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE) 
#this code makes it so we don't get annoying warning messages in our output
```


# Look at raw data

Let's load the libraries we'll need (this makes it so we can use functions like `mutate()` and `ggplot()`). We'll use **the tidyverse** for data manipulation and visualization. To access the data we will use the **fivethirtyeight** package. 


```{r load-libraries}
library(tidyverse)
library(fivethirtyeight)
```


Next we load the data. We're going to clean up the data a bit: turning the `height` variable from feet-inches to inches, getting rid of columns we don't care about, and getting rid of some columns we don't care about. 


```{r load-data}
flying_clean = 
  flying %>% 
  # turn height to inches
  mutate(height = sapply(strsplit(as.character(height),"'|\""),
        function(x){12*as.numeric(x[1]) + as.numeric(x[2])})) %>% 
  # remove variables we don't care aobut
  select(-respondent_id, -location)
```



What does this dataset look like?

```{r view-data}
flying_clean
```


Alternativey, click on `flying_clean` in the "environmnent" panel on the right to open the dataset in a new tab. What does each row represent? Each column?

To find out what the varibles mean, search for "flying" in the search box under the *help* tab in the bottom-right. 



# Analyze and visualize data


Let's look at some patterns in the data. How often do respondents say it's rude to recline your chair during the flight? (Try it out: what happens if you remove `sort = TRUE`?)


```{r recline-frequency-count}
flying_clean %>% 
  count(recline_rude, sort = TRUE)
```


Find out who people think should get the middle arm rest in a row of two seats. The `middle_arm_rest` variable contains this information. 


```{r rude-middle-arm-rest}

```


Find the count of how often people say it's rude to wake a passenger up if you are trying to go to the bathroom. The `wake_up_bathroom` variable contains this information. 


```{r rude-bathroom}

```


Find the count of how often people say they're traveling. The `frequency` variable contains this information. 


```{r frequency-bathroom}

```


## Patterns in the data


Let's make some plots to find patterns in the data. How rude do people think it is to bring an unruly baby on a plane? 

```{r, unruly-baby-count}
unruly_count = 
  flying_clean %>% 
  count(unruly_child, sort = TRUE)

ggplot(unruly_count, aes(x = unruly_child, y = n)) + 
  geom_col()
```


How rude is it to bring a baby on a plane at all? The information you want is in `baby`. 


```{r baby-count}

```


Bringing a loud baby on board is seen as a pretty rude thing to do. But do people who have children feel differently than those who don't have children? 



```{r unruly-parent}
unruly_parent = 
  flying_clean %>% 
  count(children_under_18, unruly_child)

unruly_parent

ggplot(unruly_parent, aes(x = unruly_child, y = n, fill = children_under_18)) + 
  geom_col(position = "stack") +
  labs(x = "How rude?", 
       y = "number of respondents", 
       title = "Do parents and non-parents feel differently about unruly babies?")
```


Try this: Change `position = "stack"` to `position = "dodge"`. What's different? Which plot is easier to interpret? Wny?


Instead of dodging or stacking the bars, here we'll put parents and non-parents in their own subplot, or facet. 

Is this more interpretable or less interpretable than the dodged or stacked versions from above? Change `ncol = 1` to `ncol = 2`. What's different? Which is a better plot?


```{r unruly-parents-facet}
ggplot(unruly_parent, aes(x = unruly_child, y = n, fill = children_under_18)) + 
  geom_col(position = 'stack') + 
  facet_wrap(~children_under_18, ncol = 1)
```



Let's summarize the data so we can see patterns more clearly. We'll split responses into rude ("Somewhat", "Very") and not-rude ("No"). 


```{r rude-categories}
rude_percents = 
  flying_clean %>% 
  select(unruly_child, wake_up_walk, recline_rude, wake_up_bathroom, 
         switch_seats_friends, talk_stranger, unsold_seat, baby, 
         switch_seats_family) %>% 
  drop_na() %>% 
  mutate_all(funs(ifelse(. == 'No', 0, 1))) %>% 
  summarise_all(funs(mean)) %>% 
  gather()

ggplot(rude_percents, aes(x = key, y = value)) + 
  geom_col() + 
  coord_flip() + 
  labs(y = "Percent",
       x = "Action", 
       title = "What respondents think is rude",
       subtitle = "Percent of respondents saying action is very or somewhat rude", 
       source = "FiveThirtyEight") 
  
```


What's going on in this plot? What can we learn from it?