---
title: "Problem set 7: controls"
output: html_document
---



# Instructions


- You will do your work in this `Rmarkdown` document but will submit all your answers on Canvas
- *very important*: As you move through the document, make sure to run all code chunks (grey bits) that already have code in them
- Write your own code in the empty code chunks
- Useful shortcuts: 
    - to run all the code in a specific code chunk, press the green right-facing triangle at the top right of the code chunk
    - to run all *prior* code chunks, press the downward-facing gray triangle at the top right of the code chunk
    
    
```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggdag)
library(AER)
library(huxtable)
```



# Task 1: Affairs


For this part, you'll look at data on extra-marital affairs, taken from a Psychology Today study in 1969, from the AER package. 

The affairs variable is coded weirdly, but just pretend it counts the number of affairs the respondent has had (or admits to having). 


```{r load-data}
data("Affairs")
?Affairs # read about the data and variables
head(Affairs, n = 10)
```

First, a blast from the past: find the percent of respondents who had no affairs, one affair, two affairs, and so on: 

```{r}

```


**Q1: Roughly what percent of respondents have had at least one affair?**



Now the key question: does having a child make someone more or less likely to have an extra-marital affair? 


Looking at the variable names in the Affairs dataset, make a DAG where: 

- having an affair is a function of: having children, how religious you are, how long you've been married, and how happy you are in your marriage (rating)
- having children is a function of: how long you've been married and how religious you are (more religious people tend to have more kids)
- length of marriage is a function of: how religious you are (more religious people tend to get married younger)
- happinness of marriage is a function of: whether or not you have kids (studies show having children is a significant stressor on a marriage)


```{r}


```


What variables do you need to adjust for to identify the effect of kids on affairs? 


```{r}


```


**Q2: what variables do you need to adjust for?**





Now, fit three models: 

1) a naive model, where you fail to make any adjustments. 
2) a model where you adjust for the correct variables you identified above. 
3) the same model as (2) but where you *also* control for happiness (`rating`). 


Display all three models side-by-side in a table:



```{r}



```

**Q3: How does having children affect the number of affairs someone has in the naive model?**

 


**Q4: How does having children affect the number of affairs someone has in the correct control model?**

 


**Q5: Why is the effect of having children estimated in model 3 likely to be wrong?**





# Task 2: Political advertisement


Skim this study by Huber and Arceneaux (2007): link: https://www.jstor.org/stable/4620110?seq=1#metadata_info_tab_contents

You will probably have to log-on to the school library to read it. 


The study is trying to answer an important question: can election advertisement actually persuade voters to vote for a particular candidate? Do the millions spent on ads actually work? 


Imagine you went out and collected data, across all US counties on: 1) how many dollars were spent, in ads, on pro-Biden messaging during the 2020 election; 2) how many people voted for Biden in each county. 


**Q6: Imagine you fit a model like lm(voted_biden ~ dollars_ads) with this data. What is one key reason (also noted by the the authors) for worrying about confounding in this model?**



The authors overcome these confounding concerns with a natural experiment. Thinking of this natural experiment: 

**Q7: What is the "arbitrary cutoff" the authors are taking advantage of in this natural experiment?**. 



**Q8: Explain, in your own words, why the natural experiment in this paper helps the authors identify the effect of ads on voter behavior.**



You're done! Submit the Rmd. 
