---
title: "Problem set 3: Relationships"
output: html_document
---


# Instructions


- You will do your work in this `Rmarkdown` document but will submit all your answers on Canvas
- *very important*: As you move through the document, make sure to run all code chunks (grey bits) that already have code in them
- Write your own code in the empty code chunks
- Useful shortcuts: 
    - to run all the code in a specific code chunk, press the green right-facing triangle at the top right of the code chunk
    - to run all *prior* code chunks, press the downward-facing gray triangle at the top right of the code chunk



```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(socviz)
```


# Task 1: The issues


Let's look at data on which political and social issues are most important to Americans, from a survey in 2019.  Respondents gave an answer from 1 to 4 on all of these issues, where: 

4: Very important
3: Somewhat important
2: Not very important
1: Unimportant


See the data here: 

```{r}
issues = read_csv("https://www.dropbox.com/s/x5xncajqsz0q09l/voter-files-issues.csv?dl=1")

issues
```



Calculate the *average* level of importance for all the issues:

```{r}


```
**Q1: which issue is, on average, most important to Americans?** 
**Q2: which issue is, on average, least important to Americans?** 


This is true for everyone in the sample, but what happens when we compare Democrats and Republicans? Calculate the same averages as above but *broken down by party*. 

```{r}


```

**Q3: What's the most important issue to Democrats?**: 
**Q4: What's the most important issue to Republicans?**: 


Now, I want you to compare how male and female respondents of different race groups feel about crime. Calculate the average importance of crime, broken down by race and sex. 

```{r}


```


Make a barplot using what you calculated, with race on the x-axis, average importance of crime on the y-axis, and filling the color of the bars by sex. Add labels and a theme. 

```{r}


```

**Q5: right click the graph --> "save image as" --> and save it somewhere on your computer. You will need to submit the plot!**


**Q6: Which racial group has the biggest gender gap in attitudes towards crime?** 


## Task 2: Categories


Now let's look at a survey of Americans, `gss_sm` from the `socviz` package. 

```{r}
gss_sm
```


What's the religious ID breakdown of this sample? Calculate the percent of respondents in each religious group. 

```{r}


```

**Q7: Rounding to two decimal places, what percent of the sample is Catholic?**



Next: how does each religion feel about legalizing marijuana (`grass`)? Copy your code from above and modify it so that it also tells you the percent of respondents who support and don't support marijuana legalization within each religious group. 

```{r}


```
**Q8: Which group is most pro-legalization?**




Finally: does a person's zodiac sign tell you anything about how they vote? Calculate the proportion of respondents who voted for Obama for **each** zodiac sign. Note that Obama is a dummy variable: 1, 0. 1 = voted for Obama, 0 = did not vote for Obama. 

```{r}


```
**Q9: Which zodiac sign is least likely to have voted for Obama?**


All done!
