---
title: "Data Wrangling"
citeproc: true
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-fullnote-bibliography-no-bib.csl
menu: 
  class:
    parent: Class
    weight: 3
type: docs
output:
  blogdown::html_page:
    toc: true
editor_options: 
  chunk_output_type: console
---



## In-class example


Here's the code we'll be using in class. Download it and store it with the rest of your materials for this course. If simply clicking doesn't trigger download, you should right-click and select "save link as..."

[<i class="fas fa-file-archive"></i> `wrangle-day1.R`](/slides/code/wrangle-day1.R)



## Filtering

Often, we have a big dataset that covers lots of stuff (say, all flights coming out of NYC in 2013) but we're only interested in a *subset* of those things (say, flights that arrived late over that time period). The `filter()` function is a way to subset operations that match some rule or set of rules (e.g., rule = "flights that arrived late"). We define these rules using [logical operators](https://www.statmethods.net/management/operators.html).


### Examples


Let's load the libraries.

```{r, message=F,warning=F}
# libraries
library(tidyverse)
library(nycflights13)
```


Remember you can look at the data like this.

```{r, eval=F}
# look at the data
View(flights) # open data in viewer
?flights # read data documentation
```


Let's look at flights from February.

```{r}
# look at fights, but only from February
flights %>%
  filter(month == 2)
```


Let's look at flights on Valentine's Day.

```{r}
# now let's look at flights on Valentine's Day
flights %>%
  filter(month == 2) %>%
  filter(day == 14)
```


Let's try the OR logical operator by looking at flights going to ATL or SFO.

```{r}
# try one using text and the OR symbol
# look at fights going to ATL or SFO
flights %>%
  filter(dest == "ATL" | dest == "SFO")

```


Let's look at flights between noon and 5pm.

```{r}
# try one using greater than or less than
# look at flights departing between 12pm and 5pm
flights %>%
  filter(dep_time >= 1200) %>%
  filter(dep_time <= 1700)

```

Let's look at how many flights arrived late on christmas day.

```{r}
## how many flights arrived LATE, on christmas day?
late_xmas = flights %>%
  filter(arr_time > sched_arr_time) %>%
  filter(month == 12, day == 25)

```


### Leaders


```{r}
library(juanr)
leader
```


1. A Vietnamese Emperor who, in his first year in office, was 11 years old. Famously depraved.


```{r}
leader %>%
  # first year in office
  filter(yr_office == 1) %>%
  # age at that point
  filter(age == 11) %>%
  # vietnamese
  filter(country == "VNM")
```



2. Leaders with graduate degrees who in 2015 reached their 16th year in power.

```{r}
leader %>%
  filter(edu == "Graduate", yr_office == 16, year == 2015)
```


3. The number of world leaders in the post-2000 period who have known physical or mental health issues.


```{r}
leader %>%
  filter((year > 2000) & (phys_health == 1 | mental_health == 1))
```



## Mutating

Sometimes we want to create new variables. For example, we might want to combine or alter existing variables in our dataset. The `mutate()` function is one way of doing this.


Let's convert arrival delay from minutes to hours.

```{r}
## convert arrival_delay to hours
new_flights = flights %>%
  mutate(arr_delay_hrs = arr_delay/60)
```

If you look in the dataset you will see a new variable called `arr_delay_hrs`.


Let's convert distance traveled from miles to thousands of miles.

```{r}
## convert distance to thousands of miles
new_flights2 = flights %>%
  mutate(dist_miles = distance/1000)
```


### Creating categorical variables

Sometimes we want to create more complicated variables. Here's where [case_when](https://dplyr.tidyverse.org/reference/case_when.html) comes into play.


Let's create a variable that tells us what season a flight took off in.

```{r}
## create a new variable called season
## that tells me if flight departed
## in summer, winter, fall, or spring
new_flights = flights %>%
  mutate(seasons = case_when(month == 6 ~ "Summer",
                             month == 7 ~ "Summer",
                             month == 8 ~ "Summer",
                             month == 9 ~ "Fall",
                             month == 10 ~ "Fall",
                             month == 11 ~ "Fall",
                             month == 12 ~ "Winter",
                             month == 1 ~ "Winter",
                             month == 2 ~ "Winter",
                             month == 3 ~ "Spring",
                             month == 4 ~ "Spring",
                             month == 5 ~ "Spring"))

```


We can then plot the distribution of arrival delays by season, below.

```{r}
# plot histogram of arrival delay
# separate it by season
ggplot(new_flights, aes(x = arr_delay, fill = seasons)) +
  geom_histogram() +
  facet_wrap(vars(seasons))
```


Let's say we wanted to categorize flights by how late they are. See an example, below.

```{r}
new_flights = flights %>%
  mutate(time_flight = case_when(arr_delay >= 120 ~ "very late",
                                 arr_delay > 0 & arr_delay < 120 ~ "a little late",
                                 arr_delay == 0 ~ "on time",
                                 arr_delay < 0 & arr_delay > -120 ~ "a little early",
                                 arr_delay <=-120 ~ "very early"))

```

