---
title: "Data Wrangling"
citeproc: true
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-fullnote-bibliography-no-bib.csl
menu: 
  class:
    parent: Class
    weight: 3
type: docs
output:
  blogdown::html_page:
    toc: true
editor_options: 
  chunk_output_type: console
---




## Filtering

Often, we have a big dataset that covers lots of stuff (say, all flights coming out of NYC in 2013) but we're only interested in a *subset* of those things (say, flights that arrived late over that time period). The `filter()` function is a way to subset operations that match some rule or set of rules (e.g., rule = "flights that arrived late"). We define these rules using [logical operators](https://www.statmethods.net/management/operators.html). 


### Examples


Let's load the libraries. 

```{r, message=F,warning=F}
# libraries
library(tidyverse)
library(nycflights13)
```


Remember you can look at the data like this. 

```{r, eval=F}
# look at the data
View(flights) # open data in viewer
?flights # read data documentation
```


Let's look at flights from February.

```{r}
# look at fights, but only from February
flights %>% 
  filter(month == 2)
```


Let's look at flights on Valentine's Day. 

```{r}
# now let's look at flights on Valentine's Day
flights %>% 
  filter(month == 2) %>% 
  filter(day == 14)
```


Let's try the OR logical operator by looking at flights going to ATL or SFO.

```{r}
# try one using text and the OR symbol
# look at fights going to ATL or SFO
flights %>% 
  filter(dest == "ATL" | dest == "SFO")

```


Let's look at flights between noon and 5pm.

```{r}
# try one using greater than or less than
# look at flights departing between 12pm and 5pm
flights %>% 
  filter(dep_time >= 1200) %>% 
  filter(dep_time <= 1700)

```

Let's look at how many flights arrived late on christmas day.

```{r}
## how many flights arrived LATE, on christmas day?
late_xmas = flights %>% 
  filter(arr_time > sched_arr_time) %>% 
  filter(month == 12, day == 25)

```


Let's look up the flight that Alex told us about.

```{r}
## alex search terms: 
## flight was in august
## at least 15 minutes late in departing
## headed to BOSTON
flights %>% 
  filter(dest == "BOS") %>% 
  filter(month == 8) %>% 
  filter(dep_delay >= 15)
```



## Mutating 

Sometimes we want to create new variables. For example, we might want to combine or alter existing variables in our dataset. The `mutate()` function is one way of doing this. 


Let's convert arrival delay from minutes to hours. 

```{r}
## convert arrival_delay to hours
new_flights = flights %>% 
  mutate(arr_delay_hrs = arr_delay/60)
```

If you look in the dataset you will see a new variable called `arr_delay_hrs`. 


Let's convert distance traveled from miles to thousands of miles. 

```{r}
## convert distance to thousands of miles
new_flights2 = flights %>% 
  mutate(dist_miles = distance/1000)
```


### Creating categorical variables

Sometimes we want to create more complicated variables. Here's where [case_when](https://dplyr.tidyverse.org/reference/case_when.html) comes into play. 


Let's create a variable that tells us what season a flight took off in. 

```{r}
## create a new variable called season
## that tells me if flight departed
## in summer, winter, fall, or spring
new_flights = flights %>% 
  mutate(seasons = case_when(month == 6 ~ "Summer", 
                             month == 7 ~ "Summer",
                             month == 8 ~ "Summer", 
                             month == 9 ~ "Fall", 
                             month == 10 ~ "Fall", 
                             month == 11 ~ "Fall", 
                             month == 12 ~ "Winter", 
                             month == 1 ~ "Winter", 
                             month == 2 ~ "Winter", 
                             month == 3 ~ "Spring", 
                             month == 4 ~ "Spring", 
                             month == 5 ~ "Spring"))

```


We can then plot the distribution of arrival delays by season, below. 

```{r}
# plot histogram of arrival delay
# separate it by season
ggplot(new_flights, aes(x = arr_delay, fill = seasons)) + 
  geom_histogram() + 
  facet_wrap(vars(seasons))
```


Let's say we wanted to categorize flights by how late they are. See an example, below. 

```{r}
new_flights = flights %>% 
  mutate(time_flight = case_when(arr_delay >= 120 ~ "very late",
                                 arr_delay > 0 & arr_delay < 120 ~ "a little late", 
                                 arr_delay == 0 ~ "on time", 
                                 arr_delay < 0 & arr_delay > -120 ~ "a little early", 
                                 arr_delay <=-120 ~ "very early"))

```


## Summarize


Say we want to take the average of a variable in our dataset. `summarize()` can help us do that. 


Say we wanted to know how late in departure is the *average* flight in our dataset and what's the latest a flight has ever been? 

```{r}
## on average, how late are flights in departing? 
flights %>% 
  summarise(avg_late = mean(dep_delay, na.rm = TRUE), 
            most_late = max(dep_delay, na.rm = TRUE))
```


Not the `na.rm = TRUE` above and what happens if you remove it. The problem is there are missing values (`NA`) in our data, and R can't take the average of a bunch of numbers where some are missing. `na.rm = TRUE` tells R to ignore those missing numbers and use only the complete observations. 
