---
title: "Data visualization"
citeproc: true
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-fullnote-bibliography-no-bib.csl
menu: 
  class:
    parent: Class
    weight: 2
type: docs
output:
  blogdown::html_page:
    toc: true
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```



# In-class example

Here's the code we'll be using in class. Download it and store it with the rest of your materials for this course. If simply clicking doesn't trigger download, you should right-click and select "save link as...". 

* Day one: [<i class="fas fa-file-archive"></i> `Five graphs`](/files/five-graphs.R)
* Day two: [<i class="fas fa-file-archive"></i> `Day 2`](/files/viz-day2.R)



# Making graphs with ggplot


Here we will walk through how to make some of the basic graphs in R.

The code chunk below loads our libraries and prepares the data.


```{r}
# load libraries -------------------------------------------------------------------------
library(tidyverse)
library(gapminder) # install this if you don't have it!
library(ggbeeswarm) # install this if you don't have it!
library(moderndive)

# clean data ------------------------------------------------------------------------


# subset data to focus on 2007
gap_07 =
  gapminder %>%
  filter(year == 2007)


# calculate average life span by year
life_yr =
  gapminder %>%
  select(year, lifeExp) %>%
  group_by(year) %>%
  summarise(avg_yrs = mean(lifeExp))

# calculate average life expectancy by continent
life_region =
  gap_07 %>%
  group_by(continent) %>%
  summarise(avg_yrs = mean(lifeExp))

# calculate average life expectancy by continent-year
life_region_yr =
  gapminder %>%
  group_by(continent, year) %>%
  summarise(avg_yrs = mean(lifeExp))
```



## The scatterplot


The scatterplot puts two variables on the same graph so we can see how they move together:

```{r}
# the fancy, final product
ggplot(gap_07, aes(x = gdpPercap, y = lifeExp,
                   color = continent, size = pop)) +
  geom_point()
```



We can add labels using `labs()` and a `theme` using `theme_bw()`:

```{r}
ggplot(gap_07, aes(x = gdpPercap, y = lifeExp,
                   color = continent, size = pop)) +
  geom_point() +
  labs(x = "GDP per capita ($USD, inflation-adjusted)",
       y = "Life expectancy (in years)",
       title = "Wealth and Health Around the World",
       subtitle = "Data from 2007. Source: gapminder package.") +
  theme_bw()
```


Note that there are many [more themes out there](https://ggplot2.tidyverse.org/reference/ggtheme.html).



## The time series


The time series shows you how a variable moves over time, using a line. For this one we will use the `life_yr` data object we constructed above, which looks like this:

```{r}
## look at the data
life_yr
```


We make the plot below:


```{r}
# make the plot
ggplot(life_yr, aes(x = year, y = avg_yrs)) +
  geom_line() +
  theme_bw()

```


## Multiple time series


Sometimes we have data where we observe multiple *units* moving over time, like so:

```{r}
life_region_yr
```


We can draw separate time series for each *unit* by either using the `color` aesthetic to separate the lines:


```{r}
ggplot(life_region_yr, aes(x = year, y = avg_yrs, color = continent)) +
  geom_line() +
  theme_bw()
```


Or the `group` aesthetic:


```{r}
ggplot(life_region_yr, aes(x = year, y = avg_yrs, group = continent)) +
  geom_line() +
  theme_bw()
```


## Histogram with `geom_histogram()`


```{r}
ggplot(gap_07, aes(x = lifeExp)) +
  geom_histogram() +
  theme_bw()
```


## Grouped histogram


```{r}
ggplot(gap_07, aes(x = lifeExp, fill = continent)) +
  geom_histogram() +
  theme_bw()
```


## Barplot with `geom_col()`


```{r}
ggplot(life_region_yr, aes(x = continent, y = avg_yrs)) +
  geom_col()
```


## The boxplot


```{r}
ggplot(data = gapminder, aes(x = continent, y = lifeExp)) + 
  geom_boxplot()
```





# Other layers and features


## Multi-panel plots with `facet_wrap()`

```{r}
ggplot(gapminder, aes(x = lifeExp, fill = continent)) +
  geom_histogram() +
  facet_wrap(vars(year)) +
  theme_minimal()
```



## Make aesthetics static within the geometries


```{r,echo = TRUE, out.width="60%"}
ggplot(gap_07, aes(x = gdpPercap, y = lifeExp)) + theme_light() +
  geom_point(size = 3, color = "orange", shape = 2, alpha = .5) #<<
```


Take your aesthetics out of `aes()` and into `geom()` to make them *static*



## Show a trend-line using `geom_smooth()`


```{r, echo = TRUE, out.width="60%"}
ggplot(evals, aes(x = age, y = score)) +
  geom_point() + theme_bw() + labs(x = "Professor age", y = "Student evals") +
  geom_smooth() #<<
```

Trend lines can reveal patterns in "clumpy" data


## Show *separate* trend-lines


```{r, echo = TRUE, out.width="60%"}
ggplot(evals, aes(x = age, y = score, color = gender)) + #<<
  geom_point() + theme_bw() + labs(x = "Professor age", y = "Student evals") +
  geom_smooth() #<<
```


Relationships can look different *within* groups


## Use different color and fill scales


```{r, echo = TRUE, out.width="60%"}
ggplot(gapminder, aes(x = lifeExp, fill = continent)) +
  geom_histogram() +
  scale_fill_brewer(palette = "Blues") #<<
```


`fill_brewer()` for `fill`, `color_brewer` for `color`



## Many other themes

```{r,out.width="80%"}
library(tvthemes)
ggplot(gapminder, aes(x = lifeExp, fill = continent)) +
  geom_histogram() + theme_spongeBob() + labs(title = "Horrible") +
  scale_fill_spongeBob()
```

`theme_spongeBob()` from `tvthemes` package, many more online