---
title: "Tidy data"
citeproc: true
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-fullnote-bibliography-no-bib.csl
menu: 
  class:
    parent: Class
    weight: 4
type: docs
output:
  blogdown::html_page:
    toc: true
editor_options: 
  chunk_output_type: console
---



## Tidy data


We can use the `pivot_longer` function to make data that is in "wide" format into "long" format. 


Here's an example, using the `drinks` dataset from `fivethirtyheight`. 

```{r, warning=F,message=F}
# load libraries
library(tidyverse)
library(fivethirtyeight)

# too many countries, let's look at a few
# %in% is a new logical operator: returns observations that match one of the strings
drinks_subset = 
  drinks %>% 
  filter(country %in% c("USA", "China", "Italy", "Saudi Arabia")) 


# let's gather the three alcohol variables into two: type and serving
tidy_drinks = drinks_subset %>% 
  pivot_longer(cols = c(beer_servings, spirit_servings, wine_servings), 
               names_to = "type", values_to = "serving")
tidy_drinks

# let's put position = dodge in geom_col, which will place bars side by side
ggplot(tidy_drinks, aes(x = country, y = serving, fill = type)) + 
  geom_col(position = "dodge")
```


Here's another example, using the `masculinity survey` from `fivethirtyeight`. 

```{r, warning=F,message=F}
# different dataset on masculinity
masculinity_survey


# focus on one question
# collapse age categories into long format
manly_pressure = masculinity_survey %>% 
  filter(question == "Do you think that society puts pressure on men in a way that is unhealthy or bad for them?") %>% 
  pivot_longer(names_to = "ages", 
               values_to = "percent", 
               c(age_18_34, age_35_64, age_65_over))

manly_pressure
```


And we can plot the results: 


```{r}
# plot
ggplot(manly_pressure, aes(x = response, y = percent, fill = ages)) + 
  geom_col(position = "dodge")

```


Finally, here's another example using `relig_income`. Notice here how instead of explicitly writing out every variable we want to collapse, we can just exclude the only other variable in the dataset via the "-". 

```{r}
# look at the data
relig_income

# make tidy
tidy_relig = relig_income %>% 
  pivot_longer(-religion, names_to = "income_categories", 
               values_to = "responses") %>% 
  group_by(religion) %>% 
  mutate(percent = responses/sum(responses))


# make the plot
ggplot(tidy_relig, aes(x = income_categories, y = percent)) + 
  geom_col() + 
  facet_wrap(vars(religion)) + 
  coord_flip() + 
  theme_light()
```


## Counts and percentages (group_by + tally)


Say we wanted to count how many characters in the starwars dataset have blonde, brown, etc., hair. I can do this with `group_by` and `tally`:


```{r}
starwars %>% 
  group_by(hair_color) %>% 
  tally()
```

Or, with `group_by` and `summarise` and `n()`:

```{r}
starwars %>% 
  group_by(hair_color) %>% 
  summarise(n = n())
```


Now, say I wanted to calculate the *percent* of characters with each eye color. I can do this below: 

```{r}
starwars %>% 
  group_by(hair_color) %>% 
  tally() %>% 
  mutate(percent = n/sum(n))
```


## Factor variables


Sometimes we have a categorical variable (e.g., months of the year) that we understand as having some qualitative ordering (e.g., January comes before June). R doesn't know this though, but we can tell it using `factor` variables. 


Here's an example using some data I made up: 

```{r}
# i have data on weather that looks like this:
weather = tibble(temp = c(80, 23, 14, 23, 25), 
                 month = c("January", "December", 
                           "July", "June", "October"))

weather


# i want the month variable in order
# i can use factors for this
weather_factor = weather %>% 
  mutate(month_factor = factor(month, levels = c("January", "June", 
                                          "July", "October", 
                                          "December")))
```


Notice plot without factor: 

```{r}

ggplot(weather, aes(x = month, y = temp)) + geom_col()
```

And new and imrpoved plot where month is a factor: 

```{r}
ggplot(weather_factor, aes(x = month_factor, y = temp)) + geom_col()
```


### fct_reorder

Instead of explicitly telling R how we want to order a factor, we can instead use another variable in the dataset to determine the order. 

Look at the example below, using the `starwars` dataset: 

```{r}
# starwars example
starwars

# count how many characters with each eye_color
star_eyes = starwars %>% 
  group_by(eye_color) %>% 
  tally()

star_eyes = star_eyes %>% 
  mutate(eye_color = fct_reorder(eye_color, n))

ggplot(star_eyes, aes(x = eye_color, y = n)) + 
  geom_col() + 
  coord_flip()
```

