---
title: "Correlation and regression"
citeproc: true
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-fullnote-bibliography-no-bib.csl
menu: 
  class:
    parent: Class
    weight: 5
type: docs
output:
  blogdown::html_page:
    toc: true
editor_options: 
  chunk_output_type: console
---



## The problem with objects


Most of the time in R, we are taking a data object and doing stuff to it with mutate, group_by, summarise, etc. This alters our data in different ways; to *store* these changes, we need to create a new object. See some examples below: 


```{r, eval=FALSE}
# load the libraries
library(tidyverse)
library(socviz)


# if you don't store what you've done to your data, you can't retrieve it!
gss_cat %>% 
  group_by(year) %>% 
  summarise(ave_tvhours = mean(tvhours, na.rm = TRUE))


# see? error
ggplot(gss_cat, aes(x = year, y = ave_tvhours)) + geom_col()


# you can store as a new OBJECT
tv_time = gss_cat %>% 
  group_by(year) %>% 
  summarise(ave_tvhours = mean(tvhours, na.rm = TRUE))

# or as the original object
gss_cat = gss_cat %>% 
  group_by(year) %>% 
  summarise(ave_tvhours = mean(tvhours, na.rm = TRUE))
# but note this OVERWRITES the original object

# same with mutate
gss_cat %>% 
  mutate(tv_mins = tvhours*60)

# see? error
gss_cat %>% 
  select(tv_mins)

# you can store as new object
gss_new = gss_cat %>% 
  mutate(tv_mins = tvhours*60)

# or overwrite existing (better)
gss_cat = gss_cat %>% 
  mutate(tv_mins = tvhours*60)

```


## Drawing lines (geom_smooth)


Why draw trend lines? Trend lines give us a good, educated guess as to what the value of a Y variable is given some value of X. We can draw a trend line (or line of best fit) using `geom_smooth`, as below. Notice `method = "lm"`. 


```{r, warning=F,message=F}
# libraries
library(tidyverse)
library(moderndive)

# mtcars
ggplot(mtcars, aes(x = wt, y = mpg)) + geom_point() + 
  geom_smooth(method = "lm")
```


### Why not just take an average?


Compare, for instance, against simply taking the average of every wt value. Weakness here is that we only have exact guesses for values of X for which we have data. No exact guess, for example, for a car that weighs exactly 2 tons:


```{r}
# manually
mtcars %>% 
  group_by(wt) %>% 
  summarise(ave_mpg = mean(mpg, na.rm = TRUE)) 

```


## Regression models (lm)

### continuous (number) variables

If we want to find the "math-y" parts of the line (the y-intercept, the slope), we can fit a regression model using `lm()` and interpret the output using `get_regression_table()`. 


```{r}
# finding the math-y parts of the line
# what is the y-intercept? 
# what is the slope? 
# lm(Y ~ X, data = DATA) # TEMPLATE
weight_m = lm(mpg ~ wt, data = mtcars)
get_regression_table(weight_m)
```


### categorical (word) variables

Sometimes we want to use explanatory variables that are not numbers, like car weight above, but categories (e.g., men vs. women). We can use them as below: 

```{r}
# let's look at rank
mod_rank = lm(score ~ rank, data = evals)
get_regression_table(mod_rank)
```


Interpreting is a bit tricky here. It helps to compare these coefficients to the average values of the outcome for the different categories. 

```{r}
# take the average score across ranks
evals %>% 
  group_by(rank) %>% 
  summarise(ave_score = mean(score, na.rm = TRUE))

```

Notice the correspondence: intercept above is the average eval score for teaching faculty; `ranktenure track` is the *difference* between teaching faculty and tenure track faculty (4.28 - .13 = 4.15); `ranktenured` is the *difference* between teaching faculty and tenure track faculty (4.28 - .145 = 4.14) [with some rounding]. 


The coefficients are all differences relative to the baseline. What if we want to change the baseline? Here's where factors come in. The baseline in `lm` is just the lowest factor level for the variable. So we can use `factor` to change the baseline by reordering the levels, as below :


```{r}
new_evals = 
  evals %>% 
  mutate(rank = factor(rank, levels = c("tenured", "tenure track", "teaching")))

# let's look at rank
mod_rank = lm(score ~ rank, data = new_evals)
get_regression_table(mod_rank)
```

Notice the new baseline = tenured because it's the first one on line 151. 
