---
title: "Problem set 8: Uncertainty"
output: html_document
---


# Instructions


- You will do your work in this `Rmarkdown` document but will submit all your answers on Canvas
- *very important*: As you move through the document, make sure to run all code chunks (grey bits) that already have code in them
- Write your own code in the empty code chunks
- Useful shortcuts: 
    - to run all the code in a specific code chunk, press the green right-facing triangle at the top right of the code chunk
    - to run all *prior* code chunks, press the downward-facing gray triangle at the top right of the code chunk



```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(juanr)
library(moderndive)
```





## Task 1: The draft


Imagine the following: In the 1960s all men of military age (18+ years) became draft eligible to fight in the Vietnam War. A lottery is used where the social security numbers of draft eligible men were randomly drawn. Those chosen would have to serve in the military. 


**Q1: for which of the following could the random lottery draft be used as a natural experiment?**

 - To study the effects of military service on long-term mental health
 - To study why people choose to join the military
 - To study whether drafts are effective tools of recruitment or ultimately counterproductive
 

Now imagine that, contrary to official policy, people could pay unscrupulous doctors to produce reports of illnesses and disabilities that would *disqualify* someone from the draft. 

**Q2: In which of the following ways would this fact mess up the natural experiment? Choose one** 

- The fact that many people can evade the draft will reduce the number of observations and increase our uncertainty about the effects we estimate
- Those who are drafted and not drafted are no longer as-if random, since the unscrupulous doctor is a backdoor path
- The fact would not undermine the natural experiment, since a backdoor path must affect both treatment and outcome


## Task 2: Terror


Read the abstract (and if you want, the introduction) to this paper by Rafael Di Tella and Ernesto Schargrodsky (2003), which uses a difference-in-difference design: https://bibliotecadigital.fgv.br/dspace/bitstream/handle/10438/13019/000317610_s311d.pdf?s


**Q3: What is the treatment group, control group, and the pre- and post-treatment periods in this paper?**


**Q4: What would it mean, in this paper, for the parallel trends assumption to hold? Choose one**

- The rate of Jewish institutions in the places affected and unaffected by terrorism would look similar prior to treatment onset
- The rate of robberies in places with and without Jewish institutions would look similar prior to treatment onset
- The rate of terrorist attacks in places with and without robberies would look similar prior to treatment onset


## Task 3: Feelings


```{r}
therm
?therm
```


What is the average feeling thermometer for the police in our sample (`therm`)? Calculate the average below: 

```{r}

```




How much might this estimate vary from sample to sample? Simulate this using `rep_sample_n()` by taking 1,000 samples, with replacement, each sample the same size as `therm`. Calculate the average for each of those samples and then plot the resulting averages in a histogram. 


```{r}
set.seed(1990)

```


**Q5: Between what two numbers do most of the sample averages fall?**



Repeat the same procedure above, except this time only for respondents who identify as "Middle Eastern".

```{r}
set.seed(1990)

```


**Q6: Between what two numbers do most of the sample averages fall?**



**Q7: Why is the range of values in Q5 different than in Q6?**





Look at the output from the following regression:

```{r}
lm(ft_police ~ sex, data = therm) %>% broom::tidy()
```

**Q8 Roughly speaking, what does the value in `std.error` for `sexFemale` tell us?**

- How much the effect of sex on attitudes towards the police might vary from sample to sample
- How much higher, on average, attitudes towards the police are among females than males
- The probability that females support the police more than males


**Q9 The fact that the p.value is .54 tells us that the 95% confidence interval for `sexFemale:`**

- DOES contain zero
- DOES NOT contain zero
- The p.value has no relation to confidence intervals


