---
title: "Problem set 1: Data visualization"
output: html_document
---


# Instructions


- You will do your work in this `Rmarkdown` document but will submit all your answers on Canvas
- *very important*: As you move through the document, make sure to run all code chunks (grey bits) that already have code in them
- Write your own code in the empty code chunks
- Useful shortcuts: 
    - to run all the code in a specific code chunk, press the green right-facing triangle at the top right of the code chunk
    - to run all *prior* code chunks, press the downward-facing gray triangle at the top right of the code chunk




```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(babynames) # loads babynames data, needs to install if not yet installed
library(juanr) # install if not yet installed
```


# Task 1: Baby Names


The `babynames` package includes data from the Social Security Administration with counts of baby names with at least 5 annual uses between 1880-2015.

Because there are so many names, we'll look at "Riley". Run the code below:


```{r}
baby_subset <- babynames %>%
  filter(name == "Riley")

# look at the data
head(baby_subset)
```


Using `baby_subset`, make a *time-series* that shows how often the name we picked was used for babies categorized as male and as female over time. (Hint: map the `sex` variable to the color aesthetic; also, the `n` variable shows the count per year). 


Make the plot below:


```{r}


```


**Question 1**: Roughly speaking, what happened with the use of the name Riley after the year 2000?



Now, make a similar graph for "Carey". Alter the code chunk below so that `baby_subset` contains babies named "Carey" instead of "Riley". 


```{r mckinley-subset}
baby_subset <- babynames %>%
  filter(name == "Riley")

# look at the data
head(baby_subset)
```


Now make the same plot below: 


```{r}


```


**Question 2: roughly speaking, when did the name peak in use for male and female babies?**




Finally, let's look at the most popular names in 1914, the year the first World War began. Run the code below:


```{r}
top_names = babynames %>% 
  filter(year == 1914) %>% 
  group_by(sex) %>% 
  top_n(n = 10, wt = prop) %>% 
  select(sex, name, prop) %>% 
  mutate(name = fct_reorder(name, prop))
```


Using `top_names`, make a barplot that puts the baby's name on the y-axis, and the proportion of babies with that name (the variable is called `prop`) on the x-axis. Map the sex variable to the fill aesthetic to create separate bars for male and female babies. 

```{r}


```

**Question 3: What was the 7th most popular name for male and for female babies, respectively, in 1914?**



# Task 2: The Movies


Next, let's look at some data about movies. These data are about a selection of films and info about how much they cost, how much they made in theaters, and what rating they received from reviewers. Run the code below:



```{r clean-movie-data, message=FALSE, warning=FALSE}
# read data
?movies # look at the codebook
movies = 
  movies %>% 
  # exclude films made before the 60s
  filter(year >= 1960)

head(movies)
```



Using `movies`, make a scatterplot that presents a movie's budget (x-axis) against how much it grossed in theaters (y-axis). Color the points by which `decade` the movie was made in. Add titles to the x and y axis and a theme. 

Make the plot below:


```{r budget}


```


**Question 4: right click the graph --> "save image as" --> and save it somewhere on your computer. You will need to submit the plot!**


 
Look at that outlier point in bottom-right part of the plot (should have the color for "2010s"). What movie is this? 


Make a new plot that adds labels to each of the points so that you can figure out what that movie is. Hint: look at how we did this with the health and wealth plot from class with `gapminder`.

Make the plot below (hint: if hard to see, press the icon with an arrow and a box at the top right of the image):


```{r}




```


**Question 5: What's the outlier movie?**


**Question 6: What, if anything, does the movie's position on the graph tell us about how successful it was?**




Are movies getting *longer* or *shorter* over time? Compare the distribution of movie duration across decades using a *boxplot.* 

Make the plot below: 


```{r ratings-year}


```


**Question 7: how would you describe the change in movie duration over time?**




Finally, let's look at who are the most/least *profitable* horror movie directors. First, we'll make a new dataset that only looks at scary movies, calculates each movie's profit (`gross` - `budget`) and finds the average profit for each director. To make this more manageable, let's look at directors who made at least 2 movies.


Run the code below:


```{r summarize-profits-director}
directors_profit = movies %>% 
  # only look at horror movies
  filter(genre1 == "Horror" | genre2 == "Horror" | genre3 == "Horror") %>% 
  # calculate profit
  mutate(profit = gross - budget) %>% 
  # group by director name
  group_by(director) %>% 
  # calculate number of movies each director made, average profit, average rating
  summarise(n_movies = n(), 
            avg_profit = mean(profit)) %>% 
  # include directors who made more than 2 movies
  filter(n_movies > 2)

# look at the data
head(directors_profit)
```


Now make a barplot, that has the director on the x-axis and profit on the y-axis. Hint: add `coord_flip()` to your code make the plot more readable.


Make the plot below:


```{r plot-director-profits}


```


- **Question 8: Who is the most profitable director?** 
- **Question 9: Who is the least profitable director?** 


# All done!

