---
title: "Problem set 2: Data wrangling"
output: html_document
---


# Instructions


- You will do your work in this `Rmarkdown` document but will submit all your answers on Canvas
- *very important*: As you move through the document, make sure to run all code chunks (grey bits) that already have code in them
- Write your own code in the empty code chunks
- Useful shortcuts: 
    - to run all the code in a specific code chunk, press the green right-facing triangle at the top right of the code chunk
    - to run all *prior* code chunks, press the downward-facing gray triangle at the top right of the code chunk




```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(socviz)
library(juanr)
```



# Task 1: Pokemon


First we'll look at data on pokemon. Let's load it below and look at it.


```{r load-pokemon}
pokemon
?pokemon
```


Using `filter()`, tell me which Pokemon: 

- is shorter than a meter
- weighs more than 10 times its height
- has a type 1 of "water"
- has an attack level less than or equal to 10


```{r}


```


**Q1: Who's that Pokemon?** 


Now, look at Pokemon that: 

- are taller than they are heavy
- are not legendary

```{r}


```

**Q2: What type2 do these Pokemon share in common?**



Next, look at Pokemon that are: 

- either flying or ice type1
- are better at attacking than defending
- have an HP that is greater than 40

```{r}


```

**Q3: How many Pokemon fit this description?** 



Finally, let's create some new variables and do some plotting. Construct a variable called `total_power` that is the sum of `hp`, `attack`, `defense`, and `speed`.

```{r}


```


Make a histogram plot of the distribution of total power, but facetting over the Pokemon's primary type (`type1`). Hint: look at how we used `facet_wrap` in lecture last week!

```{r}


```


**Question 4: right click the graph --> "save image as" --> and save it somewhere on your computer. You will need to submit the plot!**



## Task 2: Opiate deaths


Let's look at data on deaths related to opiate use in the US, from the `socviz` package. 


```{r}
opiates
```


First, make a barplot of opiate deaths (y-axis) by state (x-axis) but only for one year, 2010. Use `coord_flip()` so we can see the state names better. 


```{r}


```


**Question 5: What's the top state in terms of opiate deaths?**


The plot lets you compare the number of opiate deaths across states, but there's a problem: some states have a lot more people, and will therefore have *more* deaths even if the opiate crisis isn't necessarily worse there. 


To account for this, construct a new variable gives you the death *rate* per 100,000 residents. To get the rate: 

- divides deaths by population 
- multiplies that number by 100,000
- call it `deaths_per_100`


This gives you the death *rate* per 100,000 residents. Make the same barplot as above but this time with your new *rate* variable on the y-axis:

```{r}


```

**Question 6: Which state has the worst opiate death rate?**


**Question 7: Which of the following is closest to the meaning of this state having the highest opiate death rate:**

- this state has the most deaths overall in 2010
- this state has the most deaths relative to it's population in 2010
- this state has the highest year-to-year increase in opiate deaths in 2010
- the rate of opiate deaths in this state has been increasing over time 



# Task 3: World Bank categories


Let's look at data on countries' levels of national income per capita (GNI per cap):

```{r}
income = read_rds("https://github.com/hail2thief/poli-301/raw/master/static/files/gni-2021.rds") %>% 
  # drop missing income data
  drop_na(gni_cap)
income
```

The World Bank likes to categorize countries into "buckets" based on national income. The buckets are as follows: 

- Low income = less than $1,046
- Lower-middle = $1,046 – 4,095
- Upper-middle = $4,096 - 12,695
- High income = $12,695


Use `case_when` to construct a new variable called `category` that categorizes countries into these buckets. Store your new data as an object:

```{r}


```


How many countries are in the "Lower-middle" category? Replace DATA below with the name of the new data object you created above. The code will tell you how many countries are in each `category`:

```{r, eval = FALSE}
DATA %>% 
  group_by(category) %>% 
  tally()
```


**Question 8: How many countries are in the "Lower-middle" category?**



# All done!
